\documentclass[12pt, a4paper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage[bottom=3cm]{geometry}
\usepackage[hidelinks]{hyperref}
\tcbuselibrary{breakable}
\tcbuselibrary{skins}

\newtcolorbox{prob}[1]{colback=gray!5!white, colframe=gray!75!black, 
title=\textbf{Exercise #1}}

\newtcolorbox{sol}{
    breakable,
    colback=white,      % Background matches page
    colframe=white,     % Frame matches page (invisible)
    frame hidden,       % Hides the border line
    left=3mm, right=3mm,% MATCHES the Question box padding
    boxrule=0mm,        % No border width
    top=0mm, bottom=0mm,% Tight vertical spacing
    parbox=false,       % Allows paragraphs to break normally
    before upper={\textbf{Solution:}\par\medskip} % Automatically adds "Solution:" title
}

\begin{document}
\begin{prob}{2.2}
Are the random variables X and Y in the joint ensemble
of figure 2.2 independent?
\end{prob}
\begin{sol}
No, $P(x, y) \neq P(x)P(y)$ since each row or column are not proportional to each other.
\end{sol}
\bigskip

\begin{prob}{2.4}
An urn contains $K$ balls, of which $B$ are black and $W = K - B$ are white. Fred draws a ball at random from the urn and replaces
it, N times.
\begin{enumerate}
    \item[(a)] What is the probability distribution of the number of times a black
ball is drawn, $n_B$?
    \item[(b)] What is the expectation of $n_B$? What is the variance of $n_B$? What
is the standard deviation of $n_B$? Give numerical answers for the
cases $N = 5$ and $N = 400$, when $B = 2$ and $K = 10$.
\end{enumerate}
\end{prob}
\begin{sol}
\begin{enumerate}
\item[(a)] $P(n_B) = \binom{N}{n_B} (\frac{B}{K})^{n_B} (\frac{K - B}{K})^{N - n_B}$
\item[(b)] $\frac{B}{K} = \frac{1}{5}$, $\frac{K-B}{K} = \frac{4}{5}$ so the distribution is B($n$, $\frac{1}{5}$). \\[2ex]
Hence, \, $\operatorname{E}[n_b] = \frac{1}{5}n$, $\operatorname{Var}[n_B] = \frac{1}{5}n(1-\frac{1}{5})$, $\operatorname{Std}[n_B] = \sqrt{\operatorname{Var}[n_B]}$ \\[2ex]
For n = 5: \, $\operatorname{E}[n_b] = 1$, $\operatorname{Var}[n_B] = \frac{4}{5}$, $\operatorname{Std}[n_B] = \sqrt{\frac{4}{5}}$ \\[2ex]
For n = 400: \, $\operatorname{E}[n_b] = 80$, $\operatorname{Var}[n_B] = 64$, $\operatorname{Std}[n_B] = 8$ \\
\end{enumerate}
\end{sol}
\bigskip

\begin{prob}{2.5}
    An urn contains $K$ balls, of which $B$ are black and $W = K - B$ are white. We define the fraction $f_B \equiv \frac{B}{K}$. Fred draws $N$
times from the urn, exactly as in exercise 2.4, obtaining $n_B$ blacks, and
computes the quantity
\begin{equation}
z = \frac{(n_B - f_BN)^2}{Nf_B(1 - f_B)}. \tag{2.19}
\end{equation}
What is the expectation of $z$? In the case $N = 5$ and $f_B = \frac{1}{5}$, what
is the probability distribution of $z$? What is the probability that $z < 1$?
[Hint: compare $z$ with the quantities computed in the previous exercise.]
\end{prob}
\begin{sol}
\begin{align*}
    \operatorname{E}[z] &= \frac{1}{Nf_b(1-f_B)}(\operatorname{E}[n_B^2] - 2\operatorname{E}[f_B n_B N] + \operatorname{E}[f_B^2 N^2])
    &= \frac{1}{Nf(1-f)}(Nf(1-f) + N^2f^2 - 2N^2 f^2 + N^2f^2)
    &= 1
\end{align*}
With $N = 5$ and $f_B = \frac{1}{5}$,
\begin{align*}
    z &= \frac{(n_B - 1)^2}{\frac{4}{5}}
\end{align*}
so
\begin{center}
\begin{tabular}{|c|cccccc|}
    \hline
    $n_B$ & 0 & 1 & 2 & 3 & 4 & 5 \\
    \hline
    $z$ &\rule[-12pt]{0pt}{30pt} $\frac{5}{4}$ & 0 & $\frac{5}{4}$ & $5$ & $\frac{45}{4}$ & 20 \\
    \hline
\end{tabular}
\end{center}

Hence, 
$$P(z < 1) = P(n_B = 1) = \binom{5}{1} \cdot \left(\frac{1}{5}\right) \left(\frac{4}{5}\right)^4 = \frac{256}{625} = 0.4096$$
\end{sol}
\bigskip

\begin{prob}{2.8}
Assuming a uniform prior on $f_H$, $P(f_H) = 1$, solve the problem
posed in example 2.7 (p.30). Sketch the posterior distribution of $f_H$ and
compute the probability that the $N + 1$th outcome will be a head, for
\begin{enumerate}
\item[(a)] $N = 3$ and $n_H = 0$;
\item[(b)] $N = 3$ and $n_H = 2$;
\item[(c)] $N = 10$ and $n_H = 3$;
\item[(d)] $N = 300$ and $n_H = 29$.
\end{enumerate}

\end{prob}
\begin{sol}
\begin{align*}
    P(f_H | n_H, N) &= \frac{P(n_H | f_H, N) P(f_H)}{P(n_H | N)}
    &= \frac{P(n_H | f_H, N)}{P(n_H | N)} = \frac{\binom{N}{n_H} f_H^{n_H} (1-f_H)^{N-n_H}}{P(n_H | N)}
\end{align*}
Now,
\begin{align*}
    & \int P(f_H | n_H, N) df_H = 1 \\
    \iff & \frac{\binom{N}{n_H}}{P(n_H | N)} \int_{1}^{0} f_H^{n_H} (1-f_H)^{N-n_H} df_H = 1 \\
    \iff & P(n_H | N) = \binom{N}{n_H} \frac{\Gamma(n_H + 1) \Gamma(N-n_H+1)}{\Gamma(N+2)}\\
    & = \binom{N}{n_H} \frac{n_H! (N-n_H)!}{(N+1)!}\\
    & = \frac{n_H!(N-n_H)!}{(N+1)n_H!(N-n_H)!} = \frac{1}{N+1}
\end{align*}
so
$$P(f_H | n_H, N) = \frac{(N+1)!}{n_H!(N-n_H)!} f_H^{n_H} (1-f_H)^{N-n_H}$$

\vspace{1em}
The graph looks like this: \href{https://www.desmos.com/calculator/bjjlgnqg87}{www.desmos.com/calculator/bjjlgnqg87}

Now,
\begin{align*}
    \operatorname{E}[f_H] &= \int_{0}^{1} f_H P(f_H | n_H, N) df_H \\
    &= \frac{(N+1)!}{n_H!(N-n_H)!} \int_{0}^{1} f_H^{n_H+1}(1-f)^{N-n_H} df_H \\
    &= \frac{(N+1)!}{n_H!(N-n_H)!} \frac{(n_H+1)!(N-n_H)!}{(N+2)!} \\
    &= \frac{n_H+1}{N+2}
\end{align*}
Thus,
\begin{enumerate}
    \item[(a)] $\frac{1}{5}$
    \item[(b)] $\frac{3}{5}$
    \item[(c)] $\frac{4}{12} = \frac{1}{3}$
    \item[(d)] $\frac{30}{302} = \frac{15}{151}$
\end{enumerate}
\end{sol}
\bigskip

\begin{prob}{}
\end{prob}
\begin{sol}
\end{sol}
\bigskip

% template
\begin{prob}{}
\end{prob}
\begin{sol}
\end{sol}
\bigskip

\end{document}

